---
title: "How to make a submission"
date: "`r Sys.Date()`"
github_document:
    toc_dept: 1
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```


### Become a github collaborator

Please send your github user ID to oliver.ratmann@imperial.ac.uk, and accept
the invitation to collaborate on the DataChallenge.2017 repository. 


### Prepare your submission

Make your predictions in `R` and store them in a `csv` file. Submissions need to
be of the following format:

```{r, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
require(DataChallenge.2017)
data(train)
#	set the file name of your prediction
#	please replace TEAM with your team name
#	and SUBMISSION with your submission ID
outfile.base <- 'DataChallenge_TEAM_SUBMISSION_'
#	extract the countries and years for which you need to make a prediction
#	and make your prediction
sub <- subset(train, YEAR>=2015)	
sub[, PREDICTION:= as.integer(runif(nrow(sub),0,1)>0.5)]

#	your submission takes the following format: 5 columns with names
#	ISO (country code)
#	YEAR (this is obvious)
#	TEAM_ID (perhaps just your names)
#	SUBMISSION_ID (unique to this submission)
#	PREDICTION (real numbers are allowed. you should make a prediction for each country and each year)
sub		<- unique(subset(sub, select=c(ISO,YEAR,PREDICTION)))
sub[, TEAM_ID:='olli0601']
sub[, SUBMISSION_ID:='random']

#	check your submission, there should be no warnings
dc.check.submission(sub)

#	write submission to file
write.csv(sub, row.names=FALSE, file=paste0(outfile.base,'predictions.csv'))
```    

### Upload your submission

Next, you need to upload your submission. This is easy once you accepted the
github invitation to collaborate on the Data Challenge.

1. [Go to the github repository of the Data
   Challenge.](https://github.com/olli0601/DataChallenge.2017)
2. [Click on the submissions folder.](https://github.com/olli0601/DataChallenge.2017/tree/master/submissions)
3. Click the Upload Files button, drag and drop the submission you created.
4. Commit your submission to the master branch. That s it! 

### How your submission is evaluated

We will calculate the mean squared error between all of your submissions and the
true number of epidemic outbreaks in 2015 and 2016 across all countries.
Submissions that do not pass the `dc.check.submission` will be ignored. Only
your best submission will be used, so don t worry about making multiple
submissions.

Just one more note: beating our baseline prediction is not difficult, do not
worry. =J  